\bf Fast and Efficient Black Box Optimization using the Parameter-less Population Pyramid

### 2014-10-06 Fast and Efficient Black Box Optimization using the Parameter-less Population Pyramid/main.tex
Sections 6-9 reorganized.

The Random Restart Hill Climber and $1+(\lambda, \lambda)$ are both relatively effective on small problem
sizes, but have difficulty scaling to larger problems.  Only on MAX-SAT are these optimizers competitive
at larger tested problem sizes. However, this may just be an artifact as the largest tested MAX-SAT
was an order of magnitude smaller than the largest size tested for most other problems. On the Deceptive
Trap and Deceptive Step Trap problem $1+(\lambda, \lambda)$ has some other interesting results. For
problems with 4 or less traps it performs significantly better than any other algorithm. This is likely
because $1+(\lambda, \lambda)$ is able to overcome deception by probabilistically
flipping entire traps. For larger problems it becomes increasingly unlikely that mutation and crossover can
create enough changes to an individual trap to overcome deception without negatively changing others. This
behavior also causes high variance in success rate, as evident by the occasional successes on large problems.
This ability to successfully integrate modifications of more than a single bit generally appears to lead $1+(\lambda, \lambda)$
to outperform the Hill Climber on all problems except Nearest Neighbor NK.

As expected, Parameter-less hBOA was generally less efficient than hBOA using an optimal population size. The slopes
of the two lines are relatively similar across all problems, meeting with the theoretical expectation that
Parameter-less hBOA is no more than a logarithmic factor worse than hBOA.  The only problem which reversed
this trend was MAX-SAT, where Parameter-less hBOA actually outperformed hBOA. A potential explanation for
this is that hBOA is using the optimal population size for the problem class, not for each instance.
Therefore hBOA must use a population size sufficiently large to solve the hardest instances in the class.
Parameter-less hBOA on the other hand is able to solve easy instances using much smaller population sizes.
The net effect is therefore that by being more adaptable Parameter-less hBOA is able to achieve better median
results.

@@

LTGA, as the current state-of-the-art, was generally at least as fast to find the global optimum as hBOA.
However on Nearest Neighbor NK and Ising Spin Glasses LTGA appears
to perform worse than hBOA.  This may be due to LTGA's inability to represent the overlapping nature of these problems.
On MAX-SAT LTGA only outperforms hBOA and performs worse than the other 4 techniques.
Again, the likely cause to this is LTGA's requirement to have a
population size sufficiently large to solve the hardest problem in each class, reducing its effectiveness
on instances of median difficulty.

\begin{figure}
  \begin{center}
  \includegraphicsfit{evals-to-success-range}
  \end{center}
  \caption{Comparison of the upper and lower quartiles of evaluations required
           to reach the global optimum for P3 and LTGA with respect to problem size.}
  \label{fig-evals-to-success-range}
\end{figure}

Of the 130 tested configurations, P3 found the global optimum using the least median evaluations on 114.
The largest of these used 49 bits, with P3 achieving the best results on all 92 larger
configurations. hBOA, LTGA, and Parameter-less hBOA only outperform P3 on the smallest 5, 4, and 1 Deceptive
Step Trap instances, respectively. Random Restart Hill Climbing outperforms P3 on the smallest 3 Nearest Neighbor NK instances
and the smallest Ising Spin Glass. $1+(\lambda, \lambda)$ has the most success outperforming P3, doing so
on the smallest 5 Deceptive Traps, 3 smallest Deceptive Step Traps, and 2 smallest Rastrigin.

On Deceptive Trap P3 significantly outperforms both hill climbing
and LTGA by combining the best features of each, using hill climbing to optimize each trap and crossover
to perfectly mix traps. Deceptive Step Trap is less amenable to local search with P3's quality effectively
tied with both LTGA and hBOA. The slight improvement here may be due to P3's ability to solve some runs
using smaller population sizes than LTGA and hBOA due to lucky initialization. Using local search on HIFF
likely saves P3 a generation over LTGA and hBOA as all pairwise subproblems will already be solved. Combined
with P3's ability to add diversity only as needed this likely explains why P3 is able to solve HIFF a constant
multiple faster than either population based technique. P3 likely solves its median run of Nearest Neighbor
NK and Ising Spin Glasses faster than both hBOA and LTGA again by using population sizes fit to each instance
and not fit to the entire problem class. On MAX-SAT, where the most effective solvers are Hill Climber and
$1+(\lambda, \lambda)$, P3 outperforms both likely due to its use of local search with a better ability to
overcome deception.

To more fully examine the hypothesis that some of P3's success is do to its ability to scale to differences in problem difficulty,
Figure~\ref{fig-evals-to-success-range} shows the upper and lower quartiles for P3 and LTGA. On each problem
LTGA has a much smaller difference between its best and worst runs. This makes sense as LTGA uses the same
population size regardless of instance and progresses search generationally. In contrast P3 has a much
higher split, with some runs finishing very quickly. This added variability is generally on only the lower
end as P3's upper quartile is only worse than LTGA's upper quartile on Deceptive Step Trap and Nearest
Neighbor NK. A possible explanation for this is that as P3 has not pre-learned the amount of diversity required
to solve all instances does not take as risky of behavior as LTGA.

@@

% 181 total
% 126 has P3 best
% 27 LTGA
% 20 hboa
% 50 lambdalambda
% hc 9
% phboa 18

At 63\% of the recording intervals hBOA has the worst fitness of any solver. Many of the exceptions occur when
the total number of evaluations performed is significantly less than the genome size. Under these circumstances hBOA reaches better fitness
than the Hill Climber, $1+(\lambda, \lambda)$, and P3. Early in search hBOA and LTGA are both still evaluating
individuals in their randomly generated initial populations, while the three techniques worse than hBOA have started
from a single solution and are attempting to optimize. Of the methods, hBOA uses by far the largest population sizes,
about 55 times larger than LTGA on average for problems shown. As such it makes sense that hBOA would take a long time to reach high
quality solutions.

Parameter-less hBOA reported equal or higher fitness than hBOA when using the same number of total evaluations in 68\% of intervals. This makes
sense as Parameter-less hBOA's use of small populations allows it to begin to optimize long before hBOA has even evaluated
its entire initial population. However, this trend is reversed after a sufficient number of evaluations, as is apparent
on Deceptive Step Trap and Ising Spin Glasses. This makes sense as hBOA's population is set to quickly find the global
optimum.

Early in optimization P3 and the Random Restart Hill Climber have effectively identical quality. This is because
P3 performs the same evaluations as the Hill Climber for the first two restarts. Once P3 begins
performing crossover it immediately improves over the Hill Climber. In 95\% of intervals P3 had a fitness at
least as high as Hill Climbing. As such P3 is likely better than a simple
hill climber regardless of how long each technique is run and irrespective of how high quality the solution
found has to be.

Perhaps the most striking result is the quality of $1+(\lambda, \lambda)$. Until quite far into search this
method performs better than both LTGA and hBOA. Given sufficient evaluations $1+(\lambda, \lambda)$ outperforms
Random Restart Hill Climbing on all 7 problems. For brief periods in the middle of search it performs the
best of all techniques on the Deceptive Step Trap, HIFF, Ising Spin Glass, and MAX-SAT problems. A likely
explanation is $1+(\lambda, \lambda)$'s ability to efficiently incorporate gene modifications of larger
than one bit. This allows it to overcome the plateaus in Deceptive Step Trap, solve medium sized subproblems
in HIFF, flip the signs on multiple adjacent bits in Ising Spin Glass, and cross plateaus in MAX-SAT. However,
this method is not enough to quickly reach the global optima in many of these problems which causes it to eventually
be overtaken by the model building techniques.

For all problems, LTGA lags behind P3 in fitness improvement, but has a similar overall shape. On every problem except
Rastrigin, LTGA has five distinct periods: fitness plateau, near instantaneous improvement, fitness plateau,
and improvement to global optimum. The early period corresponds with
initialization of the population, with the first fitness gain achieved immediately upon completing the first generation.
This can be explained as LTGA's mixing strategy performs a sort of local search when the model has not yet detected gene epistasis.
This brings LTGA's fitness to almost the same level as the techniques explicitly performing local search, but it is not quite
as effective at reaching local optima. Subsequent generations then make minor fitness improvements, but generally stagnate
while the model remains uninformed and high quality building blocks in the population are still relatively rare. Once the model
becomes accurate enough and the probability of a crossover using high quality genetic material increases sufficiently, LTGA
enters a second period of rapid improvement.

@@

@@

It is important to note that while P3 achieved the best fitness across 70\% of time points, hBOA and LTGA were tuned specifically
to find the global optimum and are not necessarily using the optimal population size for finding intermediate fitnesses. Consider
that for both reducing the population size would almost certainly improve their fitness for evaluation steps less than their population
size. To examine the effect of population size we tested LTGA using one tenth of the population size required for reliably finding the
global optimum, with the results given in Figure~\ref{fig-small-pop}. The two problems shown are representative of the behavior of using
a smaller population size on the other five problems. In aggregate, P3 still achieved a fitness at
least as high as LTGA using the smaller population size at 80\% of intervals.
Reducing the population size resulted in LTGA's initial fitness improvement
occurring earlier, as the first generation is completed much more quickly. On Deceptive Step Trap this means that
LTGA leaps ahead of P3 due to its ability to overcome the two bit fitness plateaus. Across all tested problems using the smaller
population size reduced the quality of LTGA's first fitness plateau, likely due to missing required diversity to reach higher quality.
The second period of improvement also comes earlier and reaches a lower quality fitness. On Deceptive Step Trap 0
runs reach the global optimum, down from 100 successful for the full population size. On Nearest Neighbor NK this drops the number
of successful runs from 98 to 68. As a result we conclude that while reducing the population size of LTGA may improve its quality
at early points during optimization, doing so reduces the robustness of the final solution found. This is in contrast to P3
which balances high quality intermediate fitness without trading away eventual optimality.

\section{Computational Expenses}
\begin{figure}[t]
  \begin{centering}
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{rebuilds}
      \end{centering}
      \caption{Rebuilds}
      \label{fig-rebuilds}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{donations}
      \end{centering}
      \caption{Donations}
      \label{fig-donations}
    \end{subfigure}
  \end{centering}
  \caption{Computation costs incurred by model rebuilding (Figure~\ref{fig-rebuilds}) and
           repeated donations (Figure~\ref{fig-donations}) per evaluation as problem size increases.}
  \label{fig-costs}
\end{figure}

@@

\begin{figure}
  \begin{center}
  \includegraphicsfit{seconds-to-success}
  \end{center}
  \caption{Comparison of the median number of seconds to reach the global optimum for
           the six different optimization methods with respect
           to problem size.  If the median run did not reach the global optimum no data element
           is shown.  Results given on a log-log scale.}
  \label{fig-seconds-to-success}
\end{figure}

@@

P3 performs optimization using a mixture of hill climbing and crossover. Exactly how much time the algorithm
dedicates to each is not clear from the algorithm description alone. As such, Figure~\ref{fig-cross} shows
that the relationship between the two is strongly problem dependent but relatively independent of problem size.
In general the trend appears to suggest that problems which can be more tightly represented by crossover tend
to require less crossover evaluations. Deceptive Trap and Rastrigin spent the least evaluations on crossover, using
15\% and 20\% respectively on the largest problem sizes. Both of these problems are perfectly separable and have relatively
few local optima in each subproblem. HIFF and Ising Spin Glasses use crossover for about 60\% of evaluations. The HIFF structure can
be perfectly learned by the linkage tree, but the higher order relationships require multiple stages of crossover to optimize.
Ising Spin Glasses have an overlapping structure, making them impossible to perfectly represent in a single linkage tree, but
that overlap is relatively minor in comparison with Nearest Neighbor NK, which uses 81\% of evaluations on crossover. Deceptive Step Trap spends the highest
percentage of evaluations on crossover, 92\%. Unlike all of the other problems, the value taken by an individual bit in Deceptive
Step Trap has very little meaning. This is because any trap value is either a local optima or 1 bit flip away from being a local optima.
This has a combined result of meaning it takes very few local search evaluations to find a local optima but very many crossover operations
to proliferate the global optima. Epistatic relationships are also going to be very weak and even small population sizes are likely
to mark all clusters as $useful$.

@@

\begin{figure}[t]
  \begin{centering}
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{level-size}
      \end{centering}
      \caption{Population Size}
      \label{fig-level-size}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{level-success}
      \end{centering}
      \caption{Crossover Success}
      \label{fig-level-success}
    \end{subfigure}
  \end{centering}
  \caption{For each problem Figure~\ref{fig-level-size} shows the number of solutions stored in each level of the pyramid
           and Figure~\ref{fig-level-success} shows the percentage of fitness improving crossover evaluations at each level.}
\end{figure}

>>>>

\subsection{Quantitative Comparison}
Of the 130 tested configurations, P3 found the global optimum using the least median evaluations on 114.
The largest of these used 49 bits, with P3 achieving the best results on all 92 larger
configurations. hBOA, LTGA, and Parameter-less hBOA only outperform P3 on the smallest 5, 4, and 1 Deceptive
Step Trap instances, respectively. Random Restart Hill Climbing outperforms P3 on the smallest 3 Nearest Neighbor NK instances
and the smallest Ising Spin Glass. $1+(\lambda, \lambda)$ has the most success outperforming P3, doing so
on the smallest 5 Deceptive Traps, 3 smallest Deceptive Step Traps, and 2 smallest Rastrigin.

\subsection{Local Search}
The Random Restart Hill Climber and $1+(\lambda, \lambda)$ are both relatively effective on small problem
sizes. This is especially true for the three randomly generated problem problem classes. These problems
may contain relatively few local optima or just be exceptionally difficult for the model based algorithms.
For Deceptive Trap and Deceptive Step Trap using 4 or less traps, $1+(\lambda, \lambda)$ performs significantly
better than any other algorithm. This is likely
because $1+(\lambda, \lambda)$ is able to overcome deception by probabilistically
flipping entire traps.
This ability to successfully integrate modifications of more than a single bit generally appears to lead $1+(\lambda, \lambda)$
to outperform the Hill Climber on all problems except Nearest Neighbor NK.

On larger problem sizes, the ability for local search to reach the global optimum quickly diminishes.
Only on MAX-SAT are these optimizers competitive
at larger tested problem sizes. However, this may just be an artifact as the largest tested MAX-SAT
was an order of magnitude smaller than the largest size tested for most other problems. As the problem
size increases the number of local optima increases exponentially, which explains why Random Restart Hill
Climbing was unable to scale.
For larger problems it becomes increasingly unlikely that $1+(\lambda, \lambda)$'s mutation and crossover can
create enough changes to the correct parts of an individual required to overcome deception without negatively changing others. This
behavior also causes high variance in success rate, as evident by the occasional successes on large Deceptive Trap problems.

By integrating local search, P3 gains most of the benefits. Unlike the other model based methods, P3 generally outperforms
both Random Restart Hill Climber and $1+(\lambda, \lambda)$ on small problem sizes. On larger problem sizes
P3's ability to perform targeted crossover quickly allows it to surpass $1+(\lambda, \lambda)$'s uninformed
crossover. This targeted crossover may also explain why P3 is able to outperform $1+(\lambda, \lambda)$ on MAX-SAT.

Local search may also account for some of P3's ability to surpass LTGA and hBOA on larger problem sizes.
On Deceptive Trap P3's use of hill climbing ensures all traps are immediately optimized, allowing for perfect
linkage detection and high quality donation. Deceptive Step Trap, which is less amenable to local search, does
not show a similar improvement over the other model building techniques and performs only slightly better.
Using local search on HIFF likely saves P3 a generation over LTGA and hBOA as all pairwise subproblems will already be solved,
accounting for some of its improvement.

\subsection{Model Building}
Only techniques which explicitly built models of the search landscape were able to solve the largest problem
instances. On single instance problems LTGA was generally more effective than hBOA, with hBOA outperforming
LTGA on Nearest Neighbor NK and Ising Spin Glasses. This may be caused by the differences in modeling method:
the single instance problems can all be perfectly represented with a linkage tree while none of the randomly generated
problem classes can.

@@

Similar to hBOA and LTGA, P3's model building allows it to scale to large problem sizes. As P3 utilizes the same
model building method as LTGA, we should expect the effects of model building to be similar between those two algorithms.
Indeed on the four problems where a linkage tree can perfectly represent gene epistasis, P3 has very similar
scalability (slope) as LTGA, in most cases performing better only by a constant factor.

\subsection{Population Sizing}
\begin{figure}
  \begin{center}
  \includegraphicsfit{evals-to-success-range}
  \end{center}
  \caption{Comparison of the upper and lower quartiles of evaluations required
           to reach the global optimum for P3 and LTGA with respect to problem size.}
  \label{fig-evals-to-success-range}
\end{figure}

As expected, Parameter-less hBOA was generally less efficient than hBOA using an optimal population size. The slopes
of the two lines are relatively similar across all problems, meeting with the theoretical expectation that
Parameter-less hBOA is no more than a logarithmic factor worse than hBOA.  The only problem which reversed
this trend was MAX-SAT, where Parameter-less hBOA actually outperformed hBOA and LTGA. A potential explanation for
this is that hBOA and LTGA are using the optimal population size for the entire problem class, not for each instance.
Therefore both must use a population size sufficiently large to solve the hardest instances in the class.
Parameter-less hBOA, on the other hand, is able to solve easy instances using much smaller population sizes.
The net effect is therefore that by being more adaptable Parameter-less hBOA is able to achieve better results
on instances of median difficulty.

Scaling population size to the problem instance instead of the problem class may explain some of why P3 outperforms
LTGA and hBOA on large problem sizes. P3 solves its median run of Nearest Neighbor
NK and Ising Spin Glasses faster than both hBOA and LTGA.

To more fully examine the hypothesis that some of P3's success is do to its ability to scale to differences in problem difficulty,
Figure~\ref{fig-evals-to-success-range} shows the upper and lower quartiles for P3 and LTGA. On each problem
LTGA has a much smaller difference between its best and worst runs. This makes sense as LTGA uses the same
population size regardless of instance and progresses search generationally. In contrast P3 has a much
higher split, with some runs finishing very quickly. This added variability is generally on only the lower
end as P3's upper quartile is only worse than LTGA's upper quartile on Deceptive Step Trap and Nearest
Neighbor NK. On these problems the difference in upper quartile is very small, suggesting that even in the worst
case P3 is no worse than LTGA.

@@

% 181 total
% 121 has P3 best
%   9 hc
%  50 lambdalambda
%  20 hboa
%  18 phboa
%  27 LTGA

\subsection{Quantitative Comparison}
Of 181 sample points, P3 had the highest median fitness in 121. In pairwise competition, $1+(\lambda, \lambda)$
was the most likely to outperform P3, doing so on 50 sample points. LTGA, hBOA, and Parameter-less hBOA
were the next best, outperforming P3 on 27, 20, and 18 sample points. Random Restart Hill Climbing almost
never outperformed P3, doing so only 9 times.

\subsection{Local Search}
Perhaps the most striking result is the quality of $1+(\lambda, \lambda)$. Until quite far into search this
method performs better than both LTGA and hBOA. Given sufficient evaluations $1+(\lambda, \lambda)$ outperforms
Random Restart Hill Climbing on all 7 problems. For brief periods in the middle of search it performs the
best of all techniques on both Trap functions, HIFF, Ising Spin Glass, and MAX-SAT problems. A likely
explanation is $1+(\lambda, \lambda)$'s ability to efficiently incorporate gene modifications of larger
than one bit. This allows it to overcome the deception and plateaus in Deceptive Trap and Deceptive Step Trap, solve medium sized subproblems
in HIFF, flip the signs on multiple adjacent bits in Ising Spin Glass, and cross plateaus in MAX-SAT. However,
this method is not enough to quickly reach the global optima in many of these problems which causes it to eventually
be overtaken by the model building techniques.

The integration of hill climbing into P3 makes it strictly better than using hill climbing alone.
Early in optimization P3 and the Random Restart Hill Climber have effectively identical quality. This is because
P3 performs the same evaluations as the Hill Climber for the first two restarts. Once P3 begins
performing crossover it immediately improves over the Hill Climber. In 95\% of intervals P3 had a fitness at
least as high as Hill Climbing. As such P3 is likely better than a simple
hill climber regardless of how long each technique is run and irrespective of how high quality the solution
found has to be.

\subsection{Model Building}
Both hBOA and LTGA are marked by periods of little improvement and rapid improvement. Initially this is due to
population size effects (Section~\ref{sec-pop-sizing}). In hBOA this is taken to the extreme, with all fitness
improvement made at the very end of search. In both cases this is caused by model building. Before the model
is accurate little improvement is made. Once it is accurate, fitness improves dramatically.

On every problem LTGA has five distinct periods: fitness plateau, near instantaneous improvement, fitness plateau,
and improvement to global optimum. The early period corresponds with
initialization of the population, with the first fitness gain achieved immediately upon completing the first generation.
This can be explained as LTGA's mixing strategy performs a sort of local search when the model has not yet detected gene epistasis.
This brings LTGA's fitness to almost the same level as the techniques explicitly performing local search, but it is not quite
as effective at reaching local optima. Subsequent generations then make minor fitness improvements, but generally stagnate
while the model remains uninformed and high quality building blocks in the population are still relatively rare. Once the model
becomes accurate enough and the probability of a crossover using high quality genetic material increases sufficiently, LTGA
enters a second period of rapid improvement.

@@

\subsection{Population Sizing}
\label{sec-pop-sizing}

At 58\% of the recording intervals hBOA has the worst fitness of any solver. Many of the exceptions occur when
the total number of evaluations performed is significantly less than the genome size. Under these circumstances hBOA reaches better fitness
than the Random Restart Hill Climber, $1+(\lambda, \lambda)$, and P3. Early in search hBOA and LTGA are both still evaluating
individuals in their randomly generated initial populations, while the three techniques worse than hBOA have started
from a single solution and are attempting to optimize. Of the methods, hBOA uses by far the largest population sizes,
about 64 times larger than LTGA on average for problems shown. As such it makes sense that hBOA would take a long time to reach high
quality solutions.

Parameter-less hBOA reported equal or higher fitness than hBOA when using the same number of total evaluations in 62\% of intervals. This makes
sense as Parameter-less hBOA's use of small populations allows it to begin to optimize long before hBOA has even evaluated
its entire initial population. However, this trend is reversed after a sufficient number of evaluations, as is apparent
on Deceptive Step Trap and Ising Spin Glasses. This makes sense as hBOA's population is set to quickly find the global
optimum.

@@

It is important to note that hBOA and LTGA were tuned specifically
to find the global optimum and are not necessarily using the optimal population size for finding intermediate fitnesses. Consider
that for both reducing the population size would almost certainly improve their fitness for evaluation steps less than their population
size. To examine the effect of population size we tested LTGA using one tenth of the population size required for reliably finding the
global optimum, with the results given in Figure~\ref{fig-small-pop}. The two problems shown are representative of the behavior of using
a smaller population size on the other five problems. In aggregate, P3 still achieved a fitness at
least as high as LTGA using the smaller population size at 80\% of intervals.
Reducing the population size resulted in LTGA's initial fitness improvement
occurring earlier, as the first generation is completed much more quickly. On Deceptive Step Trap this means that
LTGA leaps ahead of P3 due to its ability to overcome the two bit fitness plateaus.
%
Across all tested problems using the smaller
population size reduced the quality of LTGA's first fitness plateau, likely due to missing required diversity to reach higher quality.
The second period of improvement also comes earlier and reaches a lower quality fitness. On Deceptive Step Trap 0
runs reach the global optimum, down from 100 successful for the full population size. On Nearest Neighbor NK this drops the number
of successful runs from 98 to 68. As a result we conclude that while reducing the population size of LTGA may improve its quality
at early points during optimization, doing so reduces the robustness of the final solution found. This is in contrast to P3
which balances high quality intermediate fitness without trading away eventual optimality.

\section{Computational Expenses}
While it is common in evolutionary computation to assume the evaluation function will dominate algorithm
complexity, in some domains this will not be true. Model based methods are especially likely to violate
this norm. Therefore in order to assess P3's quality in solving problems with efficient fitness functions,
we provide data on both its algorithmic complexity and wall clock time.

\subsection{Operation Counting}
\begin{figure}[t]
  \begin{centering}
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{rebuilds}
      \end{centering}
      \caption{Rebuilds}
      \label{fig-rebuilds}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{donations}
      \end{centering}
      \caption{Donations}
      \label{fig-donations}
    \end{subfigure}
  \end{centering}
  \caption{Computation costs incurred by model rebuilding (Figure~\ref{fig-rebuilds}) and
           repeated donations (Figure~\ref{fig-donations}) per evaluation as problem size increases.}
  \label{fig-costs}
\end{figure}

@@

\subsection{Wall Clock Performance}
\begin{figure}
  \begin{center}
  \includegraphicsfit{seconds-to-success}
  \end{center}
  \caption{Comparison of the median number of seconds to reach the global optimum for
           the six different optimization methods with respect
           to problem size.  If the median run did not reach the global optimum no data element
           is shown.  Results given on a log-log scale.}
  \label{fig-seconds-to-success}
\end{figure}

@@

\subsection{Crossover}
P3 performs optimization using a mixture of hill climbing and crossover. Exactly how much time the algorithm
dedicates to each is not clear from the algorithm description alone. As such, Figure~\ref{fig-cross} shows
that the relationship between the two is strongly problem dependent but relatively independent of problem size.
In general the trend appears to suggest that problems which can be more tightly represented by crossover tend
to require less crossover evaluations. Deceptive Trap and Rastrigin spent the least evaluations on crossover, using
15\% and 20\% respectively on the largest problem sizes. Both of these problems are perfectly separable and have relatively
few local optima in each subproblem. HIFF and Ising Spin Glasses use crossover for about 60\% of evaluations. The HIFF structure can
be perfectly learned by the linkage tree, but the higher order relationships require multiple stages of crossover to optimize.
Ising Spin Glasses have an overlapping structure, making them impossible to perfectly represent in a single linkage tree, but
that overlap is relatively minor in comparison with Nearest Neighbor NK, which uses 81\% of evaluations on crossover. Deceptive Step Trap spends the highest
percentage of evaluations on crossover, 92\%. Unlike all of the other problems, the value taken by an individual bit in Deceptive
Step Trap has very little meaning. This is because any trap value is either a local optima or 1 bit flip away from being a local optima.
This has a combined result of meaning it takes very few local search evaluations to find a local optima but very many crossover operations
to proliferate the global optima. Epistatic relationships are also going to be very weak and even small population sizes are likely
to mark all clusters as $useful$.

@@

\subsection{Pyramid}
\begin{figure}[t]
  \begin{centering}
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{level-size}
      \end{centering}
      \caption{Population Size}
      \label{fig-level-size}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
      \begin{centering}
        \includegraphicsfit{level-success}
      \end{centering}
      \caption{Crossover Success}
      \label{fig-level-success}
    \end{subfigure}
  \end{centering}
  \caption{For each problem Figure~\ref{fig-level-size} shows the number of solutions stored in each level of the pyramid
           and Figure~\ref{fig-level-success} shows the percentage of fitness improving crossover evaluations at each level.}
\end{figure}
